---
  output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# RMarkov

RMarkov is a package for computing steady-state and transient solutions of Markov chaings. Mainly, the package offers the computation funcsions for continuous-time Markov chian (CTMC).

## Installation

You can install Rsrat from github with:

```{r gh-installation, eval = FALSE}
install.packages("devtools")
devtools::install_github("okamumu/RMarkov")
```

Load the library. This package requires Matrix and deformula packages.

```{r load}
library(RMarkov)
```

## Steady-state analysis

### Steady-state solution of CTMC

Let's consider a CTMC with the following infinitesimal generator:

```{r ctmc1}
Q <- rbind(
  c(-5,2,3,0,0),
  c(1,-4,3,0,0),
  c(0,2,-3,1,0),
  c(0,0,1,-2,1),
  c(0,0,1,4,-5)
)
```

Get the steady-state solution with GTH algorithm. This is efficient to obtain the steady-state solution of a small-sized CTMC.

```{r gth}
ctmc.st.gth(Q)
```

Get the steady-state solution with GS algorithm. This is used when the CTMC is larege. Note that the matrix Q should be `CsparseMatrix (dgCMatrix)` or `dgeMatrix` when we apply `ctmc.st.gs`.

```{r gs}
ctmc.st.gs(Q)
```

### Sensitivity function of steady-state solution

The sensitivity function or high-ordered sensitivity function are defined by the first or higher derivatives of the steady-state solution. Now we have the first derivative of Q with respect to a parameter:

```{r dq}
dQ <- rbind(
  c(-1,1,0,0,0),
  c(0,-1,1,0,0),
  c(0,0,-1,1,0),
  c(0,0,0,-1,1),
  c(0,0,0,0,0)
)
```

Get the sensitivity function (the first derivative of steady-state probability)

```{r stsen}
pis <- ctmc.st.gth(Q)
ctmc.stsen.gs(Q, pis, pis %*% dQ)
```

### Quasi-stationary solution

If the underlying CTMC has absorbing states, limiting probabilities go to any of absorbing states. The quasi stationary solution is a probability vector provided that the state does not go to absorbing states even if time goes to infinity.

Let T be the infinitesimal generator over transient states for the CTMC with absorbing states.

```{r qu}
T <- rbind(
  c(-5,2,0,0,0),
  c(1,-4,3,0,0),
  c(0,2,-3,1,0),
  c(0,0,1,-2,1),
  c(0,0,0,4,-5)
)
```

The quasi stationary can be obtained by power method

```{r qs}
res <- ctmc.st.power(T)
print(res)
```

## Transient analysis

### Probability vector

Let's consider the same CTMC.

```{r ctmc2}
Q <- rbind(
  c(-5,2,3,0,0),
  c(1,-4,3,0,0),
  c(0,2,-3,1,0),
  c(0,0,1,-2,1),
  c(0,0,1,4,-5)
)
```

The initial probability vector is given by

```{r init}
x0 <- c(1,0,0,0,0)
```

Get the transient solution with uniformization.

```{r tran1}
ctmc.tran.unif(Q=Q, x=x0, t=seq(0,1,length.out = 10))
```

The reward vector is

```{r rwd}
r <- c(0,1,2,3,4)
```

```{r tran2}
res <- ctmc.tran.rwd.unif(Q=Q, x=x0, t=seq(0,20,length.out = 1000), r=r)
data <- data.frame(time=res$t, instantaneous=res$irwd, average=res$crwd / res$t)
head(data, 10)
```

```{r rwdplot1}
library(ggplot2)
data_long <- tidyr::gather(data, key="type", value=reward, -time)
ggplot(data=data_long, aes(x=time, y=reward, colour=type)) + geom_line()
```

### Sensitivity for transient solution

As similar to the previous section, we have the first derivative of Q with respect to a parameter:

```{r dq2}
dQ <- rbind(
  c(-1,1,0,0,0),
  c(0,-1,1,0,0),
  c(0,0,-1,1,0),
  c(0,0,0,-1,1),
  c(0,0,0,0,0)
)
```

The sensitivity function of transient solution can be obtained from the following expanded infinitesimal generator.

```{r transen}
Qdash <- rbind2(cbind2(Q, dQ), cbind2(RMarkov::zeroM(5), Q))
x0dash <- c(x0, rep(0,5))
rdash <- c(rep(0,5), r)
res <- ctmc.tran.rwd.unif(Q=Qdash, x=x0dash, t=seq(0,20,length.out = 1000), r=rdash)
print(res)
```

